Confirmation of \srcfull{} as a long period radio transient

### 2025-06-05 paper/main.tex
Latest synch of paper with Overleaf

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[S. J. McSweeney et al.]{
Samuel J. McSweeney$^{1}$,
Natasha Hurley-Walker$^{1}$,
Csan\'{a}d Horv\'{a}th$^{1}$,
Akash Anumarlapudi$^{2}$,
\newauthor
Angie Waszewski$^{1}$,
Dougal Dobie$^{3,4}$,
David L.~Kaplan$^{2}$,
John Morgan$^{5}$,
and Ziteng Wang$^{1}$
\\
% List of institutions
$^{1}$International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia\\
$^{2}$Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia\\
$^{3}$ARC Centre of Excellence for Gravitational Wave Discovery (OzGrav), Hawthorn, VIC 3122, Australia\\
$^{4}$Department of Physics, University of Wisconsin-Milwaukee, P.O. Box 413, Milwaukee, WI 53201, USA\\
$^{5}$CSIRO Space and Astronomy, P.O. Box 1130, Bentley, WA 6102, Australia\\
}

@@

For planning observations of \src{} at low ($\lesssim 300$\,MHz), care should be taken to incorporate scattering into the predicted ToAs.
The $N$th pulse since PEPOCH is predicted to arrive at the Solar System barycentre at
\begin{equation}
    {\rm MJD} = {\rm PEPOCH} + NP + \Delta t_{\rm DM}(\nu) + \Delta t_{\rm sc}(\nu),
    \label{eqn:timing}
\end{equation}
where PEPOCH and the period, $P$, are given in \Tab~\ref{tbl:ephemeris},
\begin{equation}
    \Delta t_{\rm DM}(\nu) = 4148\,{\rm s} \, \left(\frac{\rm DM}{{\rm pc\,cm}^{-3}}\right) \left(\frac{\nu}{\rm MHz}\right)^{-2}
\end{equation}
is the usual dispersion attributed to the interstellar medium, and
\begin{equation}
    \Delta t_{\rm sc}(\nu) = -\sqrt{2}\sigma\erfcx^{-1}\left(\frac{\tau(\nu)}{\sigma}\sqrt{\frac{2}{\pi}}\right) + \frac{\sigma^2}{\tau(\nu)}
    \label{eqn:scattering_delay}
\end{equation}
is the amount that the \emph{peak} of the observed pulse is delayed by scattering.
When calculating this quantity, we recommend using $\sigma = 15\,$s, a typical value for low-frequency pulses (see \Fig~\ref{fig:pulse_details}).
At frequencies above $300\,$MHz, $\tau(\nu) \ll \sigma$, and the delay due to scattering, which asymptotically behaves like $\Delta t_{\rm sc}(\nu) \sim \tau$, can be safely neglected.
At frequencies below $150\,$MHz, the asymptotic behaviour
\begin{equation}
    \Delta t_{\rm sc}(\nu) \sim \sigma \sqrt{\ln \left(\frac{\tau^2}{2\pi \sigma^2}\right)} + \frac{\sigma^2}{\tau}
\end{equation}
is accurate to within a few seconds.
At intermediate frequencies ($150\,{\rm MHz} \lesssim \nu \lesssim 300\,{\rm MHz}$), \Eqn~\ref{eqn:scattering_delay} should be used directly.

@@

Given \src{}'s rotation measure (RM) of ${\sim}961\,{\rm rad}/{\rm m}^2$ \citepalias[as reported in][]{2024MNRAS.535..909D}, linearly polarised emission would undergo several tens of degrees of rotation per 40\,kHz channel of the GPM observations, and therefore be somewhat depolarised.
To offer the best chance of detecting polarised emission, the MWA DDT observations between MJD 60592 and 60602 (see \Tab~\ref{tbl:obs}) were taken with a frequency resolution of 10\,kHz, which would result in only $12^\circ$ or rotation per channel at 200\,MHz.
We report no significant linearly polarised emission after de-Faraday rotating\footnote{On account of the possible sign ambiguity of the reported RM, we also tried de-Faraday rotating the MWA observations at $961\,{\rm rad}/{\rm m}^2$.} these observations at $-961\,{\rm rad}/{\rm m}^2$.
We also report that no significantly circularly polarised emission was seen in any MWA observation.

@@

As noted in \citetalias{2024MNRAS.535..909D}, optical follow-up of \src{} is hampered by its position within the Galactic plane ($b = -0.12^\circ$) and the associated extinction.
Inferring this nature of the system (white dwarf vs neutron star, isolated vs binary) from the radio pulses alone is challenging, until either the population of LPTs is better understood as a whole, or some other property of the radio signal (e.g. intermittency, polarisation) can be linked to the system properties.

@@

Although there is as yet no clear consensus on whether LPTs represent one class or multiple classes of system \citep[e.g.][]{2024ApJ...961..214R}, the recent confirmation of J1912$-$4410 as a pre-evolved white dwarf pulsar \citep{2023NatAs...7..931P} and ILT\,J1101+5521 and GLEAM-X\,J0704$-$37 as likely candidates for polars or polar-like systems \citep{deRuiter2025,2025A&A...695L...8R} raise the intriguing possibility that other LPTs may also be radio-emitting white dwarf pulsars at various stages of evolution (\citealt{2021NatAs...5..648S}, Horv\'{a}th et al. in prep).

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[S. J. McSweeney et al.]{
Samuel J. McSweeney$^{1}$,
Natasha Hurley-Walker$^{1}$,
Csan\'{a}d Horv\'{a}th$^{1}$,
Akash Anumarlapudi$^{2}$,
\newauthor
Angie Waszewski$^{1}$,
Dougal Dobie$^{3,4}$,
David L.~Kaplan$^{2}$,
John Morgan$^{5}$,
and Ziteng Wang$^{1}$
\\
% List of institutions
$^{1}$International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia\\
$^{2}$Department of Physics, University of Wisconsin-Milwaukee, P.O. Box 413, Milwaukee, WI 53201, USA\\
$^{3}$Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia\\
$^{4}$ARC Centre of Excellence for Gravitational Wave Discovery (OzGrav), Hawthorn, VIC 3122, Australia\\
$^{5}$CSIRO Space and Astronomy, P.O. Box 1130, Bentley, WA 6102, Australia\\
}

@@

For planning observations of \src{} at low frequencies ($\lesssim 300$\,MHz), care should be taken to incorporate scattering into the predicted ToAs.
The $N$th pulse since PEPOCH is predicted to arrive at the Solar System barycentre at
\begin{equation}
    {\rm MJD} = {\rm PEPOCH} + NP + \Delta t_{\rm DM}(\nu) + \Delta t_{\rm sc}(\nu),
    \label{eqn:timing}
\end{equation}
where PEPOCH and the period, $P$, are given in \Tab~\ref{tbl:ephemeris},
\begin{equation}
    \Delta t_{\rm DM}(\nu) = 4148\,{\rm s} \, \left(\frac{\rm DM}{{\rm pc\,cm}^{-3}}\right) \left(\frac{\nu}{\rm MHz}\right)^{-2}
\end{equation}
is the usual dispersion attributed to the interstellar medium, and
\begin{equation}
    \Delta t_{\rm sc}(\nu) = -\sqrt{2}\sigma\erfcx^{-1}\left(\frac{\tau(\nu)}{\sigma}\sqrt{\frac{2}{\pi}}\right) + \frac{\sigma^2}{\tau(\nu)}
    \label{eqn:scattering_delay}
\end{equation}
is the amount that the \emph{peak} of the observed pulse is delayed by scattering.
When calculating this quantity, we recommend using $\sigma = 15\,$s, a typical value for low-frequency pulses (see \Fig~\ref{fig:pulse_details}).
At frequencies above $300\,$MHz, $\tau(\nu) \ll \sigma$, and the delay due to scattering, which asymptotically behaves like $\Delta t_{\rm sc}(\nu) \sim \tau$, can be safely neglected.
At frequencies below $150\,$MHz, the asymptotic behaviour
\begin{equation}
    \Delta t_{\rm sc}(\nu) \sim \sigma \sqrt{\ln \left(\frac{\tau^2}{2\pi \sigma^2}\right)} + \frac{\sigma^2}{\tau}
\end{equation}
is accurate to within a few seconds.
At intermediate frequencies ($150\,{\rm MHz} \lesssim \nu \lesssim 300\,{\rm MHz}$), \Eqn~\ref{eqn:scattering_delay} should be used directly.

@@

Given \src{}'s rotation measure (RM) of ${\sim}961\,{\rm rad}/{\rm m}^2$ \citepalias[as reported in][]{2024MNRAS.535..909D}, linearly polarised emission would undergo several tens of degrees of rotation per 40\,kHz channel of the GPM observations, and therefore be somewhat depolarised.
To offer the best chance of detecting polarised emission, the MWA DDT observations between MJD 60592 and 60602 (see \Tab~\ref{tbl:obs}) were taken with a frequency resolution of 10\,kHz, which would result in only $12^\circ$ of rotation per channel at 200\,MHz.
We report no significant linearly polarised emission after de-Faraday rotating\footnote{On account of the possible sign ambiguity of the reported RM, we also tried de-Faraday rotating the MWA observations at $961\,{\rm rad}/{\rm m}^2$.} these observations at $-961\,{\rm rad}/{\rm m}^2$.
We also report that no significantly circularly polarised emission was seen in any MWA observation.

@@

As noted in \citetalias{2024MNRAS.535..909D}, optical follow-up of \src{} is hampered by its position within the Galactic plane ($b = -0.12^\circ$) and the associated extinction.
Inferring the nature of this system (white dwarf vs neutron star, isolated vs binary) from the radio pulses alone is challenging, until either the population of LPTs is better understood as a whole, or some other property of the radio signal (e.g. intermittency, polarisation) can be linked to the system properties.

@@

Although there is as yet no clear consensus on whether LPTs represent one class or multiple classes of system \citep[e.g.][]{2024ApJ...961..214R}, the recent confirmation of J1912$-$4410 as a pre-evolved white dwarf pulsar \citep{2023NatAs...7..931P} and ILT\,J1101+5521 and GLEAM-X\,J0704$-$37 as likely candidates for polars or polar-like systems \citep{deRuiter2025,2025A&A...695L...8R} raise the intriguing possibility that other LPTs may also be radio-emitting white dwarf pulsars at various stages of evolution (\citealt{2021NatAs...5..648S}, Horv\'{a}th et al. submitted).
